---
title: "Practical Machine Learning Course Project"
author: "jn"
date: "April 7, 2016"
output: html_document
---


##Introduction

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website
here: http://groupware.les.inf.puc-rio.br/har (http://groupware.les.inf.puc-rio.br/har) (see the section on the Weight Lifting Exercise Dataset).

This project attempts to use such data to predict what exercises people are performing.

##Source
 http://groupware.les.inf.puc-rio.br/har
```{r,message=FALSE}
#libraries
library(caret)
library(randomForest)
library(e1071)
```


##Data Loading:

```{r,cache=TRUE,message=FALSE}
setwd("~/Documents/learning/Practical ML/Course Project")
pml.training <- read.csv("pml-training.csv",na.strings = c("","NA"))
pml.testing <- read.csv("pml-testing.csv",na.strings = c("","NA"))
```

##Training and Testing Sets
Generate a testing set and a training set for the training data provided:
```{r}
# create testing set
set.seed(101)
inTrain = createDataPartition(pml.training$classe, p = 3/4)[[1]]
training = pml.training[inTrain, ]
testing = pml.training[-inTrain, ]
```

##Preliminary Data Analysis:
Gain general understanding of the dataset to develop a strategy for predictor selection. I removed descriptive columns, variables with mostly/all null values and near zero variance fields. This brings the set of columns in the training set down from 160 to 53.
```{r,message=FALSE}
#Eheck out fields
ncol(training)
#there are 160 columns
#names(training)
#Output not included in report due to length
#Fields 1:7 are identifier fields, 8:159 potential predictive/measured fields and 160 is the "classes" we are trying to predict
#Removing identifier fields leaves 152 columns.
training <- training[,8:160]
#remove near zero variance fields:
nzv <- nearZeroVar(training,saveMetrics=TRUE)
nzv.list <- nearZeroVar(training)
training <- training[,-nzv.list]
ncol(training)
# columns w/ many NAs
na.count <-sapply(training, function(x) sum(length(which(is.na(x)))))
#na.count 
#By looking at na.count, columns either have no NAs or mostly NAs.
#Remove columns with NAs
col.na <- (names(na.count[na.count > 0]))
training <- training[,!names(training) %in% col.na]
ncol(training)
#This leaves us with 53 columns
```

##Generate model
Given many variables and little understanding, and a substantial number of rows, random forest and gradient boosting models would be appropriate. Due to time and processing restraints (and high validated accuracy to be seen in the cross-validation section), I chose to use a traditional random forest method from the randomForest package.
```{r, cache=TRUE}
rf.model <- randomForest(classe ~., data=training, ntrees = 10)
```

##Analyze accuracy of model on training set
There is a 100% accuracy on the training set, but this may be due to overfitting.

```{r}
rf.train <- predict(rf.model, training)
confusionMatrix(training$classe,rf.train)
```

##Cross Validation: Analyze accuracy of model on testing set
On the testing set, there is still a 99.47% accuracy, which indicates that the methodology for determining workout positions is useful.  

```{r}
rf.train <- predict(rf.model, testing)
confusionMatrix(testing$classe,rf.train)
```



